\begin{center}
	
	\textbf{Acknowledgements}
\end{center}

\normalsize


Thanks are due to my supervisors Prof. Uwe Aickelin and Dr. Christian Wagner for their continued guidance and advice. They have indeed changed my perspective on the methods of conducting researches and publishing, showing most patience with me to gain the required level of competence in the domain of study.

I would also like to thank Prof. Simon Gaechter and Dr Felix Kolle from the School of Economics as they provide me with the required data sets of public goods game and spend the time to discuss players' behaviour and their classes.

Other members of the School of Computer Science have provided me with the required advice especially Amir Pourabdollah and Peer-Olaf Siebers. I would like also to thank all members of IMA group for providing me with motivation and friendly advice especially Shabbar Naqvi, Ian Dent, Jabran Aladi and Tuong Vu.

My thanks with respect to my father Abdulhamid Fattah and my mother Samira Bilal for their continued support and their prayers for me. This thesis would not happen if they did not culture in me the dream of persuading science and truth.  I would not forget the support which I have received from my father and mother in laws (Fawzi Abduljabbar and Adiba Bilal) and their words of motivation. I would love to thank my kids Elaf, Sidra and Mustafa for their understanding and their patience as being distant from them if not physically but certainly emotionally. Last but not the least I have reserved the warmest thanks to my lovely wife Sanar Fawzi for standing with me during this journey (As she calls it 'A chapter of life').