
\begin{center}
	\textbf{Dissemination}
\end{center}



\textbf{Journals and Chapters}

\begin{itemize}
\item Fattah, P., Aickelin, U., Wagner, C., 2016. Optimising Rule-Based Classification in Temporal Data. Zanco J. Pure Appl. Sci. 28, 135--146.

\item Fattah, P., Aickelin, U., and Wagner, C. (2017). Measuring Behavioural Change of Players in Public Goods Game. In Lecture Notes in Networks and Systems (Vol. 15).

\end{itemize}




\textbf{Conferences}
\begin{itemize}
\item Fattah, P., Aickelin, U., Wagner, C., 2016. Optimising Rule-Based Classification in Temporal Data. In: 1st International Conference on Engineering and Innovative Technology. Salahaddin University-Erbil, Erbil, Kurdistan.

\item Fattah, P., Aickelin, U., Wagner, C., 2016. Measuring Player ' s Behaviour Change over Time in Public Goods Game. In: SAI Intelligent Systems. London, UK, pp. 637--643.

\item Fatah, P., Hamarash, I., 2015. Optimization of association rule mining A two step breakdown variation of Appriori algorithm. In: Internet Technologies and Applications (ITA), 2015. IEEE, pp. 275--280.
\end{itemize}

\textbf{Presentations}

\begin{itemize}
\item Fattah, P., Aickelin, U., Wagner, C., 2014. Measuring players' behavioural change in public goods game using clustering. Network Integrated Behavioural Science.

\item Fattah, P., Aickelin, U., Wagner, C., 2014. Measuring Micro Changes over Time in Clustering. In: IMA Seminars. Nottingham.
\end{itemize}

\textbf{Posters}

\begin{itemize}
\item Fattah, P., Aickelin, U., Wagner, C., 2014. Measuring Change of Human Behaviour in Public Good Experiment. In: Kurdistan Students Conference - 2014. Nottingham.

\item Fattah, P., Aickelin, U., Wagner, C., 2013. Clustering Human Behaviour in Public Good Experiments. In: IMA 2013. Nottingham.
\end{itemize}
